---
title: "Task 5 Analysis"
output: html_notebook
---

## Load in packages

```{r}
library(naniar)

library(tidyverse)

library(janitor)

library(readr)

library(here)

library(stringr)

library(purrr)
```


## Sourcing cleaning script

```{r}
source(here("cleaning_script_task_5.R"))
```

## Reading in clean data .csv file

```{r}
read_csv(here("clean_data/rwa_clean.csv"))
```


# Analysis Questions

## What’s the average RWA score for each gender?


```{r}
rwa_clean %>% 
  group_by(gender) %>% 
  summarise(average_rwa = mean(rwa_score, na.rm = TRUE))
```
The average score for females was 2.47, for males it was 3.12, and for those
who put their gender as other the average score was 1.89


## What’s the average RWA score for left handed people vs. right handed people.


```{r}
rwa_clean %>% 
  group_by(hand) %>% 
  summarise(average_rwa = mean(rwa_score, na.rm = TRUE))
```

For those who use both hands the average score was 3.18. And for those who are 
right or left handed the average score was 2.86.


## What’s the average family size for each type of childhood?

```{r}
rwa_clean %>% 
  group_by(urban) %>% 
  summarise(average_family_size = median(familysize, na.rm = TRUE))



rwa_clean %>% 
  group_by(urban) %>% 
  summarise(average_family_size = mean(familysize, na.rm = TRUE))
```

For those who grew up in rural environments the median family size was 3 and the
mean family size was 2.79. 
For those who grew up in suburban environments the median family
size was 2 and the mean family size was 2.62.
For those who grew up in urban environments the median family size was 2 and the
mean family size was 2.50.


## What’s the average time to take the test for each education level?

```{r}
rwa_clean %>% 
  group_by(education) %>% 
  summarise(average_test_time = mean(testelapse, na.rm = TRUE))
```

For those with a graduate degree the average test time was 330 seconds.
For those with a university degree the average time taken was 623 seconds.
For those with whose highest education level was finishing high school the 
average time taken was 512 seconds.
And for those who did not finish high school the average time taken was 380
seconds.



## Create a plot of results of question 4.

```{r}
education_level_time_graph <- c(380, 512, 623, 330)

barplot(education_level_time_graph, xlab = "Education Level",
        ylab = "Seconds taken on test",
        names.arg = c("Less than high school",
                      "High School",
                      "University Degree",
                      "Graduate Degree"))
```



## What’s the average RWA score for people aged
## - Under 18
## - 18 to 25
## - 26 to 40
## - 41 to 60
## - Over 60

```{r}
rwa_age_groups <- rwa_clean %>% 
  mutate(age_group = case_when(
    age < 18 ~ "Under 18",
    age <= 25 ~ "18 to 25",
    age <= 40 ~ "26 to 40",
    age <= 60 ~ "41 to 60",
    age > 60 ~ "Over 60"
    
  ))
```


```{r}
rwa_age_groups %>% 
  group_by(age_group) %>% 
  summarise(avg_score = mean(rwa_score))
```

For those under 18 the average score was 3.36.
For those between 18 and 25 the average score was 3.15.
For those between 26 and 40 the average score was 2.61.
for these between 41 and 60 the average score was 2.54.
And for those over 60 the average scorw was 2.40.







