---
title: "R Notebook"
output: html_notebook
---

## Loading in packages

```{r}
library(naniar)

library(tidyverse)

library(janitor)

library(readr)

library(here)

library(stringr)

library(purrr)

```



## Sourcing cleaning script

In this cleaning script I cleaned the names of the raw data, then I removed 
unnecessary columns. After that I removed the duplicate columns. Then I removed
the "£" from the food spend column and made it numeric. Finally I made both the 
sizes and gender of the dogs uniform in the way they were written, but still
distinguishable from one another. I then wrote this as a .csv in the clean_data
folder.

```{r}
source(here("cleaning_script_task_6.R"))
```


## Reading clean .csv file

```{r}
dogs_clean <-  read_csv(here("clean_data/dogs_survey_clean.csv"))

```


# Analysis Question


## The client only counts a valid email address as one ending in ‘.com’. How 
## many survey results have a valid email address.
 
```{r}
dogs_clean %>% 
  filter(str_detect(email, pattern = "\\.com$"))
```

181 clients had valid email addresses


## What’s the average amount spent on dog food for each dog size?

```{r}
dogs_clean %>% 
  group_by(dog_size, na.rm = TRUE) %>% 
  summarise(avg_food_spend = mean(food_spend, na.rm = TRUE))
```

For XS dogs the average spend was £57.30, for S dogs £51.51, for M dogs the
average spend was £47.26, for L dogs the average spend was £51.84, and for XL
dogs the average spend was £51.07.


## For owners whose surname starts with a letter in the second half of the 
alphabet (N onwards) what is the average age of their dog? 

```{r}
dogs_clean %>% 
  filter(str_detect(last_name, "^(?i)[N-Z]")) %>% 
  summarise(avg_age = mean(dog_age, na.rm = TRUE))
  
```

The average age for dogs whose owners whose surnames start with letters in the 
second half of the alphabet (N-Z) was 53.3 dog years.


## The dog_age column is the age in dog years. If the conversion is 1 
## human year = 6 dog years, then what is the average human age for dogs of each 
## gender?

```{r}
dogs_clean %>% 
  group_by(dog_gender, na.rm = TRUE) %>% 
  summarise(avg_human_age = (mean(dog_age, na.rm = TRUE)/6))
```

The average age for female dogs was 8.0 human years, and for male dogs it was 
8.5 human years


## Create a plot of results of question 4.

```{r}
dog_age_graph <- c(8.0, 8.5)

barplot(dog_age_graph, 
        xlab = "Dog Gender",
        ylab = "Age in Human Years",
        names.arg = c("Female",
                      "Male"))
```



# Summary of Analysis

I first looked to see how many of the clients had valid email addresses for the
system, which was determined by the email ending in ".com". From the data I 
could see that 181 clients had valid email addresses.

Then I looked to see what the average amount spent on dog food was, depending
on which size category the dog fell under. From the data I could see that for 
XS dogs the average spend was £57.30, for S dogs £51.51, for M dogs the
average spend was £47.26, for L dogs the average spend was £51.84, and for XL
dogs the average spend was £51.07.

After I looked to what the average age in dog years was for clients whose last
name started with letters which fell under the second half of the alphabet
(N-Z). I found that the average age for dogs whose owners whose surnames start 
with letters in the second half of the alphabet (N-Z) was 53.3 dog years.

I finally looked into what the average age in human years of the dogs were 
accoridng to their gender. I found that he average age for female dogs was 8.0 
human years, and for male dogs it was 8.5 human years. I also did a plot of this
data.









