---
title: "Task 1 Analysis"
output: html_notebook
---
#Loading in packages

```{r}
library(naniar)

library(tidyverse)

library(janitor)

library(readr)

library(here)

library(stringr)
```

#Sourcing cleaning script

```{r}
source(here("cleaning_script_task_1.R"))
```


#Reading in clean_data .csv file
```{r}
read_csv(here("clean_data/decathlon_clean.csv"))
```


#Who had the longest long jump seen in the data?

```{r}
decathlon_clean %>% 
  select(name, long_jump) %>% 
  slice_max(long_jump, with_ties = TRUE)
```
Clay had the longest long jump at 7.96m


#What was the average 100m time in each competition?

```{r}
decathlon_clean %>% 
  summarise(average_100m = mean(`100m`))
```
Average 100 meter time was 11.0


#Who had the highest total points across both competitions?

```{r}
decathlon_clean %>% 
  group_by(name) %>% 
  summarise(total_score = sum(points)) %>% 
  slice_max(total_score, with_ties = TRUE)
```
Sebrle had the highest score across both competitions with 17110 points


#What was the shot-put scores for the top three competitors in each competition?

```{r}
decathlon_clean %>% 
  group_by(competition) %>% 
  slice_max(shot_put, n=3, with_ties = TRUE)
```

In the decastar competition the top three scores for shot-put were: 15.19, 
14.83, and 14.77

In the OlympicG the top three shot-put scores were: 16.36, 15.93, and 15.73


#What was the average points for competitors who ran the 400m in less than 50 
seconds vs. those than ran 400m in more than 50 seconds?

```{r}
decathlon_clean %>% 
  mutate(
    less_than_50 = case_when(
      `400m` < 50 ~ TRUE,
      `400m` > 50 ~ FALSE
  )) %>% 
  group_by(less_than_50) %>% 
  summarise(average_points = mean(points))
```

For those that ran the 400m in less than 50 seconds the average points were 8120
And for those that ran the 400m in over 50 seconds the average points were 7727
