---
title: "Cake Analysis"
output: html_notebook
---

## Loading in packages

```{r}
library(naniar)

library(tidyverse)

library(janitor)

library(readr)

library(here)

library(stringr)
```


## Sourcing cleaning script

This .R script was used to take in the raw data, clean the column names to make 
them uniform in case. I then cleaned the data to make it appropriate for use in 
analysis. I did this by pivoting the data to make it long format, I joined the two tables to get the full ingredient names, I then selected the appropriate columns and dropped any columns with NAs as they were irrelevant to the analysis. I also made the columns all lowercase to keep the data uniform. 
I then wrote the cleaned data as a .csv file into the "clean_data" folder

```{r}
source(here("cleaning_script_task_2.R"))
```


## Reading in clean cake data

```{r}
read_csv(here("clean_data/cake_clean.csv"))
```


# Analysis Questions

## Which cake has the most cocoa in it?

```{r}
cake_clean %>% 
  filter(ingredient == "cocoa") %>% 
  slice_max(amount, with_ties = TRUE)
```

One Bowl Chocolate cake had the most cocoa in it with 10 tablespoons


## For sponge cake, how many cups of ingredients are used in total?

```{r}
cake_clean %>% 
  filter(cake == "sponge",
         measure == "cup") %>% 
  summarise(total_cups = sum(amount))
```

In total 3.5 cups of ingredients were used


## How many ingredients are measured in teaspoons?

```{r}
cake_clean %>% 
  filter(measure == "teaspoon") %>% 
  summarise(teaspoon_ingredients = n_distinct(ingredient))
```

8 different ingredients are measured using teaspoons

## Which cake has the most unique ingredients?

```{r}
cake_clean %>% 
  group_by(cake) %>% 
  summarise(n_ingredients = n()) %>% 
  slice_max(n_ingredients, with_ties = TRUE)
```

Both babas au rhum cake, and one bowl chocolate cake have 11 ingredients


## Which ingredients are used only once?

```{r}
cake_clean %>% 
  group_by(ingredient) %>% 
  summarise(n_uses = n()) %>% 
  filter(n_uses == 1)
```

The ingredients that were only used once were: bananas, cream of tartar, crushed
ice, dried currants, egg white, nutmeg, nuts, and zweibach.


# Analysis summary

Using the cake_clean data I first found that the cake with the most cocoa in it
was the one bowl chocolate cake with 10 tablespoons.

I then looked to see how many cups of ingredients were used to make the sponge
cake and found that 3.5 cups of ingredients were used.

I then found that in total 8 different ingredients in the data were measured 
using tablespoons

I looked to see which cakes used the most ingredients and found that both babas 
au rhum cake, and one bowl chocolate cake have 11 ingredients

Finally I looked into which ingredients were only used once across all the 
different cakes and found that the ingredients that were only used once were: 
bananas, cream of tartar, crushed ice, dried currants, egg white, nutmeg, nuts, 
and zweibach.








