---
title: "Birds Analysis"
output: html_notebook
---

#Reading in clean data csv

```{r}
here()

ships_birds_clean <- read_csv(here("clean_data/ships_birds_clean.csv"))
  
```


#Which bird had the most individual sightings?

```{r}
ships_birds_clean %>% 
  group_by(species_abbreviation) %>% 
  summarise(individual_sightings = n_distinct(record_id)) %>% 
  arrange(desc(individual_sightings)) %>% 
  slice_max(individual_sightings, n = 1, with_ties = TRUE) %>% 
  pull(species_abbreviation)
  
```

DIOANTEXU was the most common species abbreviation, with 5632 individual 
sightings. This abbreviation is representative of the "Wandering Albatross"



#Which bird had the highest total count?


```{r}
ships_birds_clean %>% 
  group_by(species_abbreviation) %>% 
  summarise(total_count = sum(count)) %>% 
  arrange(desc(total_count)) %>% 
  slice_max(total_count, n = 1, with_ties = TRUE) %>% 
  pull(species_abbreviation)
```

PUFTEN was the species abbreviation with the highest total count of sightings at 
982553 total sightings. 
PUFTEN is representative of the short-tailed shearwater species.


#Which bird had the highest total count above a latitude of -30?

```{r}
ships_birds_clean %>% 
  filter(lat > -30) %>% 
  group_by(species_abbreviation) %>% 
  summarise(total_count = sum(count)) %>% 
  arrange(desc(total_count)) %>% 
  slice_max(total_count, n = 1, with_ties = TRUE) 
```

The wedge-tailed shearwater was the bird with the highest count above a 
latitiude of -30 with a total count of 855





#How many different types of birds were only ever seen in groups of 1?

```{r}

ships_birds_clean %>% 
  summarise(individual_species = n_distinct(species_abbreviation))


ships_birds_count_1 <- ships_birds_clean %>% 
  filter(count == 1)


ships_birds_greater_1 <- ships_birds_clean %>% 
  filter(count > 1)


ships_birds_count_1 %>% 
  anti_join(ships_birds_greater_1, by = "species_abbreviation") %>% 
  summarise(solo_species = n_distinct(species_abbreviation)) %>% 
  pull()


```

In total there were 23 distinct species that were only ever seen in groups of 
one


# How many penguins were seen? 

```{r}
ships_birds_clean %>% 
  filter(
      str_detect(species_common_name_taxon_age_sex_plumage_phase, 
                 "(?i)penguin") == TRUE) %>% 
  summarise(total_count = sum(count)) %>% 
  pull()

```

In total 158 penguins were observed











