---
title: "Birds Analysis"
output: html_notebook
---

#Reading in clean data csv

```{r}
ships_birds_clean <- read_csv(here("clean_data/ships_birds_clean.csv"))
  
```


#Which bird had the most individual sightings?

```{r}
ships_birds_clean %>% 
  group_by(species_abbreviation) %>% 
  summarise(individual_sightings = n_distinct(record_id)) %>% 
  arrange(desc(individual_sightings)) %>% 
  slice_max(individual_sightings, n = 1, with_ties = TRUE) %>% 
  pull(species_abbreviation)
  
```

DIOANTEXU was the most common species abbreviation, with 5632 individual 
sightings. This abbreviation is representative of the "Wandering Albatross"



#Which bird had the highest total count?


```{r}
ships_birds_clean %>% 
  group_by(species_abbreviation) %>% 
  summarise(total_count = sum(count)) %>% 
  arrange(desc(total_count)) %>% 
  slice_max(total_count, n = 1, with_ties = TRUE) %>% 
  pull(species_abbreviation)
```

PUFTEN was the species abbreviation with the highest total count of sightings at 
982553 total sightings. 
PUFTEN is representative of the short-tailed shearwater species.


#Which bird had the highest total count above a latitude of -30?

```{r}
ships_birds_clean %>% 
  filter(lat > -30) %>% 
  group_by(species_abbreviation) %>% 
  summarise(total_count = sum(count)) %>% 
  arrange(desc(total_count)) %>% 
  slice_max(total_count, n = 1, with_ties = TRUE) 
```

The short-tailed shearwater was the species with the highest count of 
observations above a latitude of -30, in total there were 30379 of the species
spotted






