---
title: "Task 4 Analysis"
output: html_notebook
---

## Loading in packages

```{r}
library(naniar)

library(tidyverse)

library(janitor)

library(readxl)

library(here)

library(stringr)
```

## Sourcing cleaning script

This file is an R script for cleaning the three different datasets in the 
raw_data folder and bind them into one dataset to be used in analysis

Each data set was very messy and required a lot of cleaning as seen in the 
script. I had to make the columns uniform in all of them so they could be 
binded correctly. And i removed any rows where an opinion hadnt been given to a 
candy. There was some particularly messy values in the countries
column which required 'hard-coding'. When the tables were binded I removed all 
the values in the candy_type column that weren't actually candy, and i combined
certain candy types if they had been written differently across the three 
years.

```{r}
source(here("cleaning_script_task_4.R"))
```


## Reading in clean_candy.csv
```{r}
here()

clean_candy <- read_csv("clean_data/clean_candy.csv")
```



## What is the total number of candy ratings given across the three years. (Number of candy ratings, not the number of raters. Don’t count missing values)

Already removed NAs in opinions column so can just count the number of rows

```{r}
colSums(is.na(clean_candy))


clean_candy %>% 
  select(opinion) %>% 
  count()


```

589655 ratings were given across the three years


## What was the average age of people who are and are not going out trick or 
treating?

```{r}
clean_candy %>% 
  select(going_tot, age) %>% 
  group_by(going_tot) %>% 
  summarise(mean_age = mean(age)) 



clean_candy %>% 
  select(going_tot, age) %>% 
  group_by(going_tot) %>% 
  summarise(median_age = median(age))
```

The mean age of those going trick or treating was 35.4 years.
The median age was 38

The mean age of those who weren't going trick or treating is 39.3
The median age was 38


## For each of joy, despair and meh, which candy bar received the most of these ratings?

```{r}
clean_candy %>% 
  group_by(opinion, candy_type) %>% 
  summarise(count = n()) %>% 
  slice_max(count, with_ties = TRUE)
```

It was "any full sized candy bar" which received the most joy ratings
It was jolly ranchers which received the most meh ratings
It was jolly ranchers which received the most despair ratings




## How many people rated Starburst as despair?

```{r}
clean_candy %>% 
  filter(candy_type == "starburst") %>% 
  group_by(opinion) %>% 
  summarise(total_ratings = n())
```

1867 people rated starburst as DESPAIR



### What was the most popular candy bar by this rating system for each gender in the dataset ?

```{r}
clean_candy_numbers <- clean_candy %>% 
  mutate(opinion = as.numeric(
    recode(
      opinion, 
      "JOY" = 1,
      "DESPAIR" = -1,
      "MEH" = 0
    
  )))
```



```{r}
clean_candy_numbers %>% 
  group_by(gender, candy_type) %>% 
  summarise(rating = sum(opinion)) %>% 
  slice_max(rating, with_ties = TRUE)
```
For males "any full sized candy bar" was the highest rated

For females "any full sized candy bar" was the highest rated

For those who put their gender as other the highest rated candy was regular m&ms

For those who would rather not say their gender the highest rated candy was 
regular m&ms



## What was the most popular candy bar in each year?

```{r}
clean_candy_numbers %>% 
  group_by(year, candy_type) %>% 
  summarise(rating = sum(opinion)) %>% 
  slice_max(rating, with_ties = TRUE)
```
In 2015 the highest rated candy was "Any full sized candy bar"

In 2016 the most "any full sized candy bar" was regular m&ms

In 2017 the most popular candy was regular m&ms


## What was the most popular candy bar by this rating for people in US, Canada, UK, and all other countries?

```{r}
clean_candy_numbers %>% 
  group_by(country, candy_type) %>% 
  summarise(rating = sum(opinion)) %>% 
  slice_max(rating, with_ties = TRUE)
```
The most popular candy in the US was "any full sized candy bar"

The most popular candy in the UK was a five way tie between: 
"any full sized candy bar", lindt truffle, rolos, toblerone, and regular m&ms

The most popular candy in Canada was "any full sized candy bar"

The most popular candy in all the other countries together was 
"any full sized candy bar"


## Summary of Analysis

My first question was to look at how many candy ratings had been given out over 
the three years. As I had already removed any NAs from the opinion column I 
was able to simply add up how many rows were in that column and got an answer
of 589655 ratings given

I then looked at the average age of people who went trick or treating and found
a median age of 38 and a median age of 35.4.

For those who didn't go trick or treating the median age was also 38, but the 
average age was 39.3

For the opinion the distinct values in the column were "Joy", "Meh", and 
"Despair", depending on how the person filling out the table felt about each
respective candy. I looked to see which candies received the most of each of 
these ratings. For "joy" "it was "any full sized candy bar" which received the 
most, and for "despair" and "meh" it was jolly ranchers which received the most.

I looked at breaking down how many of each distinct opinion had been given for
Starburst candy, and found that 4849 people put "joy", 1046 people put "meh",
and 1867 people put "despair".

After this I created a new table which was designed to rank the candies with a 
numeric system based on the previous 'word' ranking system.

I used this system to find out how candy had been rated across the various 
genders in the table. I found that for male and female "any full sized candy 
bar" was the highest rated. And for those who put their gender as other or
"prefer not to say" regular m&ms were the highest rated.

I then used the system to see which were the most popular candies across each of
the three years in the table and found that in 2015 and in 2016 the most popular 
candy was"any full sized candy bar", and in 2016. and 2017 the most popular 
candy was regular m&ms.

Finally I broke the table down by country and found the most popular candy in
the UK was a five way tie between: "any full sized candy bar", lindt truffle, 
rolos, toblerone, and regular m&ms. And in the US, Canada, and all the other 
countries combined the most popular candy was was "any full sized candy bar".








