---
title: "R Notebook"
output: html_notebook
---
#Task 4 Analysis


```{r}
library(naniar)

library(tidyverse)

library(janitor)

library(readxl)

library(here)

library(stringr)
```





```{r}
here()

clean_candy <- read_csv("clean_data/clean_candy.csv")
```



##What is the total number of candy ratings given across the three years. (Number of candy ratings, not the number of raters. Don’t count missing values)

Already removed NAs in opinions column so can just count the number of rows

```{r}
colSums(is.na(clean_candy))



clean_candy %>% 
  select(opinion) %>% 
  count()



```

589655 ratings were given across the three years


##What was the average age of people who are going out trick or treating?

```{r}
clean_candy %>% 
  select(going_tot, age) %>% 
  filter(going_tot == "Yes") %>% 
  summarise(average_age = mean(age))



clean_candy %>% 
  select(going_tot, age) %>% 
  filter(going_tot == "Yes") %>% 
  summarise(average_age = median(age))
```

The mean age of those going trick or treating was 35.4 years.
The median age was 38


##What was the average age of people who are not going trick or treating?

```{r}
clean_candy %>% 
  select(going_tot, age) %>% 
  filter(going_tot == "No") %>% 
  summarise(average_age = mean(age))



clean_candy %>% 
  select(going_tot, age) %>% 
  filter(going_tot == "No") %>% 
  summarise(average_age = median(age))
```
The mean age of those who weren't going trick or treating is 39.3
The median age was 38


## For each of joy, despair and meh, which candy bar received the most of these ratings?

```{r}
clean_candy %>% 
  filter(opinion == "JOY") %>% 
  group_by(candy_type) %>% 
  summarise(count = n()) %>% 
  slice_max(count, with_ties = TRUE)
```

It was regular m&ms which received the most joy ratings


```{r}
clean_candy %>% 
  filter(opinion == "DESPAIR") %>% 
  group_by(candy_type) %>% 
  summarise(count = n()) %>% 
  slice_max(count, with_ties = TRUE)
```

It was jolly ranchers which received the most despair ratings

```{r}
clean_candy %>% 
  filter(opinion == "MEH") %>% 
  group_by(candy_type) %>% 
  summarise(count = n()) %>% 
  slice_max(count, with_ties = TRUE)
```

It was regular m&ms which received the most meh ratings


##How many people rated Starburst as despair?

```{r}
clean_candy %>% 
  filter(candy_type == "starburst") %>% 
  group_by(opinion) %>% 
  summarise(total_ratings = n())
```

4849 people rated starburst as JOY



###What was the most popular candy bar by this rating system for each gender in 
###the dataset ?

```{r}
clean_candy_numbers <- clean_candy %>% 
  mutate(opinion = as.numeric(
    recode(
      opinion, 
      "JOY" = 1,
      "DESPAIR" = -1,
      "MEH" = 0
    
  )))
```


```{r}
clean_candy %>% 
  distinct(gender)
```


```{r}
clean_candy_numbers %>% 
  filter(gender == "Male") %>% 
  group_by(candy_type) %>% 
  summarise(rating = sum(opinion)) %>% 
  slice_max(rating, with_ties = TRUE)
```
For males regular m&ms were the highest rated

```{r}
clean_candy_numbers %>% 
  filter(gender == "Female") %>% 
  group_by(candy_type) %>% 
  summarise(rating = sum(opinion)) %>% 
  slice_max(rating, with_ties = TRUE)
```

For females regular m&ms were the highest rated


```{r}
clean_candy_numbers %>% 
  filter(gender == "Other") %>% 
  group_by(candy_type) %>% 
  summarise(rating = sum(opinion)) %>% 
  slice_max(rating, with_ties = TRUE)
```

For those who put their gender as other the highest rated candy was regular m&ms

```{r}
clean_candy_numbers %>% 
  filter(gender == "I'd rather not say") %>% 
  group_by(candy_type) %>% 
  summarise(rating = sum(opinion)) %>% 
  slice_max(rating, with_ties = TRUE)
```
For those who would rather not say their gender the highest rated candy was 
regular m&ms



##What was the most popular candy bar in each year?

```{r}
clean_candy_numbers %>% 
  filter(year == 2015) %>% 
  group_by(candy_type) %>% 
  summarise(rating = sum(opinion)) %>% 
  slice_max(rating, with_ties = TRUE)
```
In 2015 the highest rated candy was "Any full sized candy bar"


```{r}
clean_candy_numbers %>% 
  filter(year == 2016) %>% 
  group_by(candy_type) %>% 
  summarise(rating = sum(opinion)) %>% 
  slice_max(rating, with_ties = TRUE)
```

In 2016 the most popular candy was regular m&ms

```{r}
clean_candy_numbers %>% 
  filter(year == 2017) %>% 
  group_by(candy_type) %>% 
  summarise(rating = sum(opinion)) %>% 
  slice_max(rating, with_ties = TRUE)
```
In 2017 the most popular candy was regular m&ms


##What was the most popular candy bar by this rating for people in US, Canada, UK, and all other countries?

```{r}
clean_candy_numbers %>% 
  filter(country == "US") %>% 
  group_by(candy_type) %>% 
  summarise(rating = sum(opinion)) %>% 
  slice_max(rating, with_ties = TRUE)
```
The most popular candy in the US was regular m&ms


```{r}
clean_candy_numbers %>% 
  filter(country == "UK") %>% 
  group_by(candy_type) %>% 
  summarise(rating = sum(opinion)) %>% 
  slice_max(rating, with_ties = TRUE)
```

The most popular candy in the UK was regular m&ms

```{r}
clean_candy_numbers %>% 
  filter(country == "Canada") %>% 
  group_by(candy_type) %>% 
  summarise(rating = sum(opinion)) %>% 
  slice_max(rating, with_ties = TRUE)
```

The most popular candy in Canada was regular m&ms


```{r}
clean_candy_numbers %>% 
  filter(country == "Other") %>% 
  group_by(candy_type) %>% 
  summarise(rating = sum(opinion)) %>% 
  slice_max(rating, with_ties = TRUE)
```
The most popular candy in all the other countries together was regular m&ms

